---
output:
  html_document: default
  pdf_document: default
---

Decisions (labeled by D_ in the code): 

In our data, we decided to only include students who successfully graduated from Amherst College. 
Within Amherst College, we only included people who graduated with a Bachelor of Arts 
We combined our data so that there were six data points for study, for each student: graduation date, major, major type (single vs double/triple), graduate university, grad school type, and grad school graduation date. 


```{r}
# Stat231: Clearinghouse Challenge (scratch work)
# Spring 2020
# 

# load packages here
library(tidyverse)
library(janitor)
library(lubridate)
```


```{r}
 # import data here
 # do not copy the links provided here
 # only use them to import the data while working in R on the project

 demo_data <- read.csv("https://www.dropbox.com/s/e0td697qjb3l9hc/demo%20file.csv?dl=1") %>%
   clean_names()  %>%
#rename(temp_id = i_temp_id)

 clearinghouse_data <- read.csv("https://www.dropbox.com/s/rwctyw777fodry2/002115st_398963_DETLRPT_SE_02242020121415_002115se0019.csv?dl=1") %>%
   clean_names()

cip_codes <- read.csv("https://www.dropbox.com/s/j4q7qt7di31ejfq/AC_Programs_Div_and_CIP.csv?dl=1") %>%
  clean_names()
```

``` {r}

# get acquianted with the data
# D1
modified <- clearinghouse_data %>%
  filter(graduated == "Y")%>%
  mutate(graduation_date = ymd(graduation_date))%>%
  select(graduation_date, degree_title, requester_return_field, college_name) 
#D2
clear2 <- modified %>%
  filter(college_name == "AMHERST COLLEGE" & degree_title == "BACHELOR OF ARTS")
#What about degrees from Amherst that are not BA's?

clear3 <- modified %>%
  filter(college_name != "AMHERST COLLEGE") %>%
  rename(post_BA = college_name,
         Post_BA_Graduation_Date = graduation_date)

#D3 
clear4<-full_join(clear2, clear3, by = "requester_return_field", suffix = c(" BA", " Post-BA")) %>%
  distinct(requester_return_field, .keep_all = TRUE)
 #D3 cont

demo_d<- demo_data %>%
  rename(requester_return_field = temp_id)%>%
  select(requester_return_field, student_program)
#D3 cont

clear5 <- inner_join(clear4, demo_d)
#D3 cont
clear6 <- mutate(clear5, majorType = ifelse(str_detect(clear5$student_program,"-"),"Multiple","Single"), yearsToGrad = (clear5$Post_BA_Graduation_Date-clear5$graduation_date)/365)
clear6 <- clear6  %>% 
  rename("college" =college_name) %>% 
  rename("majors" =student_program) %>% 
  rename("studentID" =requester_return_field) %>%
  rename("gradType" =`degree_title Post-BA`) %>%
  rename("gradSchool" =post_BA) %>%
  rename("gradSchool_gradDate" =Post_BA_Graduation_Date) 

arranged <- clear6 %>% select(c(3,4,8,9,1,7,6,5,10))

```


