---
title: "Stat231: Lab 13 - Unsupervised Learning" 
author: "Your Name Here"
date: "April 28, 2020"
always_allow_html: yes
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes 
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggrepel)

sessionInfo()

knitr::opts_chunk$set(eval = TRUE
                      , message = FALSE
                      , warning = FALSE
                      , fig.width = 9
                      , fig.height = 6)
```

# K-means clustering 

### SAT Scores

Let's begin with a simple example of clustering some colleges and universities based on characteristics of an incoming class of first years.  To simplify our look at this data, we will only look at a random sample of schools in Massachusetts, and we will focus on SAT scores for right now.

```{r}
colleges <- read_csv("http://kcorreia.people.amherst.edu/F1920/CollegeScoreCard1617.csv")

# names(colleges)

set.seed(20191107)
ma_samp <- colleges %>% 
  mutate(SATMT = as.numeric(SATMT25)
         , SATVR = as.numeric(SATVR25)) %>%
  filter(STABBR %in% c("MA") & !is.na(SATMT) 
            & !is.na(SATVR)) %>%
  select(STABBR, CITY, INSTNM, SATMT, SATVR) %>%
  sample_n(15)

ma_samp$INSTNM
```

First, let's just look at a scatterplot of Math SAT (25th percentile) vs Verbal SAT (25th percentile):

```{r}
ggplot(data = ma_samp, aes(x = SATMT, y = SATVR)) +
  geom_point() + 
  #geom_text(aes(label=INSTNM), nudge_y = 10) 
  geom_text_repel(aes(label=INSTNM), size = 3)
```

Which schools are most similar?

> ANSWER:Gordon College and MCLA

Do we need to standardize the variables in this case? Why or why not?

> ANSWER:No

Let's use k-means to find two clusters:

```{r}
set.seed(23)
vars <- c("SATMT", "SATVR")
km2_out <- kmeans(ma_samp[,vars], centers=2, nstart=20)
km2_out

# vector of cluster assignments
km2_out$cluster

# add cluster assignment to the data frame
ma_samp_clust3 <- ma_samp %>%
  mutate(cluster_assign = as.character(km2_out$cluster))

# the centroids for the fit
km2_out$centers

# visualize the cluster assignments and centroids
ggplot(data = ma_samp_clust3, aes(x = SATMT, y = SATVR)) + 
  geom_point(aes(color = cluster_assign)) +
  coord_fixed() +
  geom_point(data = as.data.frame(km2_out$centers)
             , aes(x = SATMT, y = SATVR)
             , pch = "X"
             , size = 4) +
  labs(x = "Math SAT 25th Percentile"
       , y = "Verbal SAT 25th Percentile" 
       , color = "Cluster Assignment")
```

What if we used 5 clusters?  Repeat the k-means clustering analysis and visualization above, but with 5 clusters instead of 2.  

```{r}
km5_out <- kmeans(ma_samp[,vars], centers=5, nstart=20)

# add cluster assignment to the data frame
ma_samp_clust5 <- ma_samp %>%
  mutate(cluster_assign = as.character(km5_out$cluster))

# visualize the cluster assignments and centroids
ggplot(data = ma_samp_clust5, aes(x = SATMT, y = SATVR)) + 
  geom_point(aes(color = cluster_assign)) +
  coord_fixed() +
  geom_point(data = as.data.frame(km5_out$centers)
             , aes(x = SATMT, y = SATVR)
             , pch = "X"
             , size = 4) +
  labs(x = "Math SAT 25th Percentile"
       , y = "Verbal SAT 25th Percentile" 
       , color = "Cluster Assignment")

```

Which solution will have *smaller* total within-cluster variation?  Why?  Check your answer using the code below.

> ANSWER: The five clusters, as they are smaller more grouped togther clusters with a smaller range. 

```{r}
km2_out$tot.withinss
km5_out$tot.withinss
```


#### Plotting $W_k$ vs. K

```{r}
fig <- matrix(NA, nrow=10, ncol=2)

set.seed(75)
for (i in 1:10){
  fig[i,1] <- i
  fig[i,2] <- kmeans(ma_samp[,4:5]
                    , centers=i
                    , nstart=20)$tot.withinss
}

ggplot(data = as.data.frame(fig), aes(x = V1, y = V2)) +
  geom_point() + 
  geom_line() +
  labs(x = "K", y = expression("W"[k]))
```

### Standardization

Let's try clustering colleges by more than two variables.  This time, we'll consider the following set of characteristics:

- Admission rate
- SAT average
- Median debt at graduation
- % of graduates majoring in Mathematics & Statistics
- Average cost of attendance

This will be harder to visualize, but since we only have 15 observations, it's not hard to visually inspect which colleges end up in each cluster.

```{r}
# add the additional variables needed to our sample dataset 
ma_samp2 <- colleges %>%
  select(INSTNM, ADM_RATE, SAT_AVG
         , GRAD_DEBT_MDN, PCIP27, COSTT4_A) %>%
  right_join(ma_samp, by = "INSTNM") %>%
  mutate(adm_rate = as.numeric(ADM_RATE)
         , sat_avg = as.numeric(SAT_AVG)
         , grad_debt_mdn = as.numeric(GRAD_DEBT_MDN)
         , avg_cost = as.numeric(COSTT4_A)
         , prop_math = as.numeric(PCIP27)
         # standardize variables
         , adm_rate_norm = scale(adm_rate)
         , sat_avg_norm = scale(sat_avg)
         , grad_debt_norm = scale(grad_debt_mdn)
         , avg_cost_norm = scale(avg_cost)
         , prop_math_norm = scale(prop_math)) %>%
  # kmeans() function will return error if any variables missing
  filter(!is.na(adm_rate) & !is.na(sat_avg) 
         & !is.na(grad_debt_mdn) & !is.na(avg_cost) 
         & !is.na(prop_math))

# k=3 
set.seed(23)

## on standardized variables
vars_norm <- c("adm_rate_norm", "sat_avg_norm", "grad_debt_norm", "avg_cost_norm", "prop_math_norm")
km3_out <- kmeans(ma_samp2[,vars_norm], centers=3, nstart=20)

## on unstandardized variables (for comparison)
vars2 <- c("adm_rate", "sat_avg", "grad_debt_mdn", "avg_cost", "prop_math")
km3_out_un <- kmeans(ma_samp2[,vars2], centers=3, nstart=20)


# add cluster assignments to the data frame
ma_samp2_clust3 <- ma_samp2 %>%
  mutate(cluster_assign_norm = as.character(km3_out$cluster)
         ,cluster_assign = as.character(km3_out_un$cluster))
```

What schools are clustered together when using the standardized variables?  Are the clusters different if the unstandardized variables are used?

> ANSWER: They do seem different. Cluster 2 is a lot bigger in the standard assignment whereas cluster 3 is larger in the undstandard.


```{r}
ma_samp2_clust3 %>%
  arrange(cluster_assign_norm) %>%
  select(INSTNM, cluster_assign_norm)

ma_samp2_clust3 %>%
  arrange(cluster_assign) %>%
  select(INSTNM, cluster_assign)

```


\newpage
# Hierarchical clustering

### SAT Scores

Again working with the `ma_samp` dataset, compute the distance between SAT scores (25th percentile for Math & Verbal) for all 105 *pairs* of schools.  Note that we do not have to standardize the variables here because they are similar in size/range. 

```{r}
# how many pairs of schools are there?  
choose(15,2)
15*14/2

vars
d <- dist(ma_samp[,vars])
class(d)
d

# computation for [2,1]: 86.02325
d[1]
# distance between 1st and 2nd obs
ma_samp[1:2,c("INSTNM",vars)]
sqrt((470-420)^2 + (490-420)^2)
```

Now we can use these distances to get hierachical clustering:

```{r}
ma_hc <- hclust(d)
ma_hc[["height"]]
```

Next we can plot the dendogram:

```{r}
plot(ma_hc, labels = ma_samp$INSTNM)
```

We can cut the tree to get a specific number of clusters, or we can cut the tree at a specific height. First, let's cut the tree to get 3 clusters.

```{r}
cutree(ma_hc, k=3)

# add classifications to data frame
ma_samp3 <- ma_samp %>%
  mutate(hc_clust3 = as.character(cutree(ma_hc, k=3)))

# plot clusters
ggplot(data = ma_samp3, aes(x = SATMT, y = SATVR
                            , color = hc_clust3)) +
  geom_point() +
  geom_text_repel(data = filter(ma_samp3, INSTNM == "Amherst College")
            , aes(label=INSTNM)) +
  labs(x = "Math SAT 25th Percentile"
       , y = "Verbal SAT 25th Percentile" 
       , color = "Cluster Assignment")
```

What if we cut the tree at a height of 70?  How many clusters are there?

> ANSWER: It looks like 5 clusters,but I'm not entirely sure. 

```{r}
ma_hc[["height"]]

# visualize on dendogram
plot(ma_hc, labels=ma_samp$INSTNM)
abline(h=70, lty=2)

# cut the tree at a specific height
cutree(ma_hc, h=70)

# add classifications to data frame
ma_samp4 <- ma_samp %>%
  mutate(hc_clust5 = as.character(cutree(ma_hc, h=70)))

# plot clusters
ggplot(data = ma_samp4, aes(x = SATMT, y = SATVR
                            , color = hc_clust5)) +
  geom_point() +
  geom_text_repel(data = filter(ma_samp4, INSTNM == "Amherst College")
            , aes(label=INSTNM)) +
  labs(x = "Math SAT 25th Percentile"
       , y = "Verbal SAT 25th Percentile" 
       , color = "Cluster Assignment")
```

### Standardization

Now fit a hierarchical clustering model considering the following set of characteristics:

- Admission rate
- SAT average
- Median debt at graduation
- % of graduates majoring in Mathematics & Statistics
- Average cost of attendance 

(Use `ma_samp2` dataset.)  Make a basic dendogram with label names.

```{r}


```


Which school seems most similar to Gordon College?  Can we tell which school seems most similar to Amherst College? Compare some of the variable values to verify your answer.

> ANSWER:

```{r}



```


What clusters would you get if you cut the tree at a height of 3?  Try to identify the answer by looking at the dendogram; then check your resonse with the `cutree` function.  (Note: You may need to knit the PDF and look at the dendogram there in order to really be able to see it ...)

> ANSWER:

```{r}


```

We can add cluster colors to the dendrogram using the `ColorDendrogram` function in the `sparcl` package.  Let's load that library (install it first, if needed), and create a colored dendogram based on three clusters. 

```{r, eval=FALSE}
library(sparcl)

ColorDendrogram(ma_hc2, y = cutree(ma_hc2, k=3)
                , labels = ma_samp2$INSTNM
                , branchlength = 1.6)
```

What are the characteristics of each of these clusters? To visualize this, we need to look at the variable values for schools within each group. Here we create a narrow verison of the data to facet by variable and compute density curves for each cluster group:

```{r, eval=FALSE}
ma_samp5 <- ma_samp2 %>% 
  select(adm_rate, sat_avg, grad_debt_mdn, prop_math, avg_cost) %>%
  mutate(cluster_hc3 = as.character(cutree(ma_hc2, k=3)))

ma_samp5 %>%
  gather(key=variable, value=value, -cluster_hc3) %>%
  ggplot(aes(x=value, color=cluster_hc3)) + 
  geom_density() + 
  facet_wrap(~variable, scales="free") 
```


Clustering methods don't just determine clusters based on individual variable values, but how these variables combine with all other variables (e.g. clusters in $R^5$ space for this example). We can also look at a 2-d view of the data using a scatterplot matrix. We can use the ggplot "add on" package called `GGally` to do this:

```{r}
library(GGally)

ggpairs(data = ma_samp5, aes(color = cluster_hc3)
          , columns = vars2)
```

Look at the scatterplot for median graduation debt and proportion math graduates.  Why might there be a green school within a cluster of red schools?

> ANSWER:



# Your choice

Choose a *larger* subgroup of colleges and universities that might be of interest to you (e.g., all institutions in Massachusetts; or all institutions in New England; or all . . .).  Explore the data using k-means clustering, hierarchical clustering, and/or PCA.  You could stick to using the variables we've considered already, or you could consider other variables.  The data dictionary for the College Scorecard data has been posted on Moodle, and can help you identify what the other variables are, if interested.

> ANSWER:

```{r}

```

# References

Data is from: https://collegescorecard.ed.gov/data/

