---
title: "STAT 231: Problem Set 3"
author: "Xander Schwartz"
date: "due by 5 PM on Monday, February 24"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

This homework assignment is designed to help you futher ingest, practice, and expand upon the material covered in class over the past week(s).  You are encouraged to work with other students, but all code and text must be written by you, and you must indicate below who you discussed the assignment with (if anyone).  

Steps to proceed:

\begin{enumerate}
\item In RStudio, go to File > Open Project, navigate to the folder with the course-content repo, select the course-content project (course-content.Rproj), and click "Open" 
\item Pull the course-content repo (e.g. using the blue-ish down arrow in the Git tab in upper right window)
\item Copy ps3.Rmd from the course repo to your repo (see page 6 of the GitHub Classroom Guide for Stat231 if needed)
\item Close out of the course repo project in RStudio
\item Open up your repo project in RStudio
\item In the ps3.Rmd file in YOUR repo, replace "YOUR NAME HERE" with your name
\item Add in your responses, committing and pushing to YOUR repo in appropriate places along the way
\item Run "Knit PDF" 
\item Upload the pdf to Gradescope
\end{enumerate}

```{r, setup, include=FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
```


\newpage 
## If you discussed this assignment with any of your peers, please list who here:

> ANSWER:

\newpage
# MDSR Exercise 5.4

In the `Marriage` data set included in the `mosaicData` package, the `appdate`, `ceremonydate`, and `dob` variables are encoded as factors, even though they are dates.  Convert the three columns into a date format.  After converting the variables, use `glimpse()` to output a glimpse of the dataset.

```{r}
library(mosaicData)
library(lubridate)
library(Lahman)

Marriage  <- Marriage %>% mutate(appdate = mdy_hms(appdate))%>% mutate(ceremonydate = mdy_hms(ceremonydate))%>% mutate(dob = mdy_hms(dob))

Marriage%>%select(appdate,ceremonydate,dob)%>%glimpse()

```


\newpage
# MDSR Exercise 5.5

Consider the values returned by the `as.numeric()` and `readr::parse_number()` functions when applied to the following vectors (`x1` and `x2`).  Describe the results and their implications.

> ANSWER: As numeric is only able to read 1900.45 correctly, as the other numbers include a non-numeric symbol. As numeric registers x2 as some sort of whole number, as the factor version screws it up. Parse number is far more succesfsul  of the x1, but for x2 it does not work, as x2 as already been set to a factor. 

```{r}
x1 <- c("1900.45", "$1900.45", "1,900.45", "nearly $2000")
x2 <- as.factor(x1)



as.numeric(x1)

as.numeric(x2)

parse_number(x1)
parse_number(x2)
```


\newpage
# Algorithmic thinking: gathering and spreading without `gather()` and `spread()`

This exercise is designed to both deepen your understanding of how the `gather` and `spread` functions work "behind the scenes", and to practice algorithmic thinking -- how would one go about transposing a dataset from narrow to wide format or wide to narrow format if the `gather()` and `spread()` functions didn't exist?   

a. Using the `Teams` dataset from the `Lahman` package, create a dataset called `Teams_Wide` that includes one row per team and contains the following variables withOUT using the `spread()` function:

- `lgID` (League ID)
- `teamID` (Team ID)
- `W_1988` (the number of wins a team had in 1988)
- `W_1998` (the number of wins a team had in 1998)
- `W_2008` (the number of wins a team had in 2008)
- `W_2018` (the number of wins a team had in 2018)

You cannot use the `spread()` function here, but you should be able to create this dataset using only: `filter`, `select`, `rename`, and `full_join`. 

```{r}

# Teams_Temp <- Teams %>% filter(yearID == 1988|yearID == 2008|yearID == 2018|yearID == 1998)%>%select(yearID,lgID,teamID)
# Teams_Temp2 <- Teams %>% filter(yearID == 1988|yearID == 2008|yearID == 2018|yearID == 1998)%>%select(yearID,W)
# Teams_Wide <- full_join(Teams_Temp,Teams_Temp2)
Teams88 <- Teams %>% filter(yearID == 1988)%>%select(teamID,lgID,yearID,W)%>%rename(c("W"="W_1988"))
Teams98 <- Teams %>% filter(yearID == 1998)%>%select(teamID,W)%>%rename(c("W"="W_1998"))
Teams08 <- Teams %>% filter(yearID == 2008)%>%select(teamID,W)%>%rename(c("W"="W_2008"))
Teams18 <- Teams %>% filter(yearID == 2018)%>%select(teamID,W)%>%rename(c("W"="W_2018"))

Teams_Wide <- full_join(Teams88,Teams98,by="teamID")
Teams_Wide <- full_join(Teams_Wide,Teams08,by="teamID")
Teams_Wide <- full_join(Teams_Wide,Teams18,by="teamID")%>%select(-yearID)



```


b. Now create a dataset called `Teams_Wide2` that is the same as `Teams_Wide` but is created using the `spread()` function. 

```{r}
FilteredTeams <- Teams %>% filter(yearID == 1988|yearID == 2008|yearID == 2018|yearID == 1998)%>%select(yearID,lgID,teamID,W)
Teams_Wide2 <- spread(FilteredTeams, key= yearID, value = W)
```

c. Next, *working from `Teams_Wide2`*, create a dataset called `Teams_Narrow` withOUT using `gather()` that contains the following variables:

- `lgID` (League ID)
- `teamID` (Team ID)
- `year` (Year)
- `wins` (the number of wins a team had in a given year)

You cannot use the `gather()` function here, but you should be able to create this dataset using only: `mutate`, `select`, `rename`, and `bind_rows`. 

```{r}
# teamsFirst <- Teams_Wide[1:2]
# teamsTwo <- Teams_Wide[3]
# teamsTest <- bind_rows(teamsFirst,teamsTwo)
# teamsTwo <- Teams_Wide[4]
# teamsTest <- bind_rows(teamsTest,teamsTwo)
# teamsTest <- bind_rows(teamsTest[1:3],teamsTest[4])

teamsHold  <- Teams_Wide %>%  bind_rows(Teams_Wide[3],Teams_Wide[4],.id = "Year98")
teamsHold2  <- teamsHold %>%  bind_rows(teamsHold[4],teamsHold[6],.id = "Year08")
teamsHold3  <- teamsHold2 %>%  bind_rows(teamsHold2[5],teamsHold2[8],.id = "Year18")%>%select(c(1:6))
Teams_Narrow <- teamsHold3  %>% mutate(Year = ifelse(Year18==1&Year08==1&Year98==1,1998, ifelse(Year18==1&Year08==2,2008,ifelse(Year18==1&Year08==3,2018,1988))))%>%select(c(4:7))






```

d. Lastly, create a dataset (starting from from `Teams_Wide`) called `Teams_Narrow2` that is the same as `Teams_Narrow` but is created using the `gather()` function. 

```{r}

Teams_Narrow2<-  gather(Teams_Wide, key  = "year", value = "wins",W_1988:W_2018)
```

\newpage
# User-defined functions

I don't want to discourage great questions in class by adding more to the homework when a great question is asked . . . so, instead of adding a problem to update the `count_name()` function we discussed in class, I will instead allow you to choose whether you want to answer an original problem I had written for this problem set (writing a function related to MDSR Figure 3.22) OR updating the `count_names()` function.  (Note that I think the original problem is more difficult, so you can thank the student who spurred this other question which has allowed you to opt to solve an easier problem :) )

Choose ONE of (a) or (b) to answer:

## a. Age Distribution of American ? Named ?

Re-create Figure 3.22 from MDSR (page 53) of the "Age Distribution of American Boys Named Joseph".  Then, write a function called `names_plot` that takes inputs of `nm` (for name) and `sx` (for sex) that would reproduce the figure, but for different names/sexes.  Be sure that the figure includes an arrow pointing to the highlighted bar, and includes all of the text as included in Figure 3.22 (e.g. "Number of [name]s born each each", "The median living [name] is [?] years old"), but updated to correspond to that name and sex.  

You can set the x-axis to always be from 1990 to 2010 (the range of the data), but the y-axis should change depending on the name/sex in order for the plot to be most visible/effective. 

Once your function is written, apply the function for:

- `names_plot(nm = "Donald", sex = "M")`
- `names_plot(nm = "Elizabeth", sex = "F")`
- `names_plot(nm = "Bernie", sex = "M")`
- `names_plot(nm = "Pete", sex = "M")`
- `names_plot(nm = "Amy", sex = "F")`

```{r}
library(babynames)
library(mdsr)
BabynamesDist <- make_babynames_dist()

nmplot <- ggplot(data = BabynamesDist %>% filter (name == nm & sex = sex),  aes (x=year)) + geom_bar(stat="identity", aes(y = count_thousands * alive_prob),fill="#b2d7e9",colour = "white") + geom_line (aes(y=count_thousands),size=2) +  ylab ("Number of poeple (Thousands)") + xlab(NULL) 

wtd.quantile <- Hmisc::wtd.quantile
median_yob <- 
  with(joseph, wtd.quantile(year, est_alive_today, probs = 0.5))
median_yob

nmplot <- nmplot +
  geom_bar(stat = "identity", colour = "white", fill = "#008fd5", 
           aes(y = ifelse(year == median_yob, est_alive_today / 1000, 0)))
 nmplot <- nmplot +
  ggtitle("Age Distribution of American Boys Named" + nm) + 
  geom_text(x = 1935, y = 40, label = "Number of" + nm + "nborn each year") + 
  geom_text(x = 1915, y = 13, label = 
    "Number of" +  nm + "\nborn each year\nestimated to be alive\non 1/1/2014", 
    colour = "#b2d7e9") + 
  geom_text(x = 2003, y = 40, 
    label = "The median\nliving "+ nm + "\nis"  + median_yob + "years old", 
            colour = "darkgray") + 
  geom_curve(x = 1995, xend = 1974, y = 40, yend = 24, 
    arrow = arrow(length = unit(0.3,"cm")), curvature = 0.5) + ylim(0, 42)


namePlot <- function(nm, sex)
{
  nmplot  %+% filter(BabynamesDist, name == nm, sex == sex)

}

namePlot("Donald", "M")


```


## b. Updating `count_name()`

Update the `count_name()` function defined in Lab05a to allow the `nm` input to be a vector of names up to length 20 (i.e. the vector can be anywhere from one name to twenty names).  (More than 20 names would make for an illegible figure.)  Have the function output a plot of year (x-axis) and total number of births (y-axis), with a title that includes all the names in the `nm` vector.  The function should return a relevant, informative error message if more than 20 names are given in the vector or none of the given names have a match.  (Hint: using the `stop()` function to catch errors early on can yield a more efficient function.)

Test your function on the following inputs and be sure they return the correct plot or error message:

- your name and two friends' names in proper case (e.g. Katharine)
- your name and two friends' names in all lower case letters (e.g. katharine) 
- our class (`count_names(our_class)` with `our_class` defined below)


```{r, error = TRUE}

our_class <- str_to_title(c("aaron", "ainsley", "aleksandar", "beth"
           , "bobby", "brandon", "breanna", "chris"
           , "courtney", "dario", "dickson", "grace"
           , "isabelle", "jacob", "janet", "jordan"
           , "konstantin", "maggie", "matt", "michael"
           , "natacha", "stephanie", "umaimah", "xander"))

```


\newpage
# Trump Tweets

David Robinson, Chief Data Scientist at DataCamp, wrote a blog post "Text analysis of Trump's tweets confirms he writes only the (angrier) Android half"[http://varianceexplained.org/r/trump-tweets/].

He provides a dataset with over 1,500 tweets from the account realDonaldTrump between 12/14/2015 and 8/8/2016.  We'll use this dataset to explore the tweeting behavior of realDonaldTrump during this time period.

First, read in the file. Note that there is a `TwitteR` package which provides an interface to the Twitter web API.  We'll use this R dataset David created using that package so that you don't have to set up Twitter authentication.  

```{r}
load(url("http://varianceexplained.org/files/trump_tweets_df.rda"))
```

## A little wrangling to warm-up

1a.  There are a number of variables in the dataset we won't need.  

- First, confirm that all the observations in the dataset are from the screen-name `realDonaldTrump`.  

- Then, create a new dataset called `tweets` that only includes the following variables:

- `text`
- `created`
- `statusSource`

```{r}
tweets <- select(trump_tweets_df,text,created,statusSource)
```

\newpage
1b. Using the `statusSource` variable, compute the number of tweets from each source.  How many different sources are there?  How often are each used?

> ANSWER: There are five different sources. There are 1 instagram and twitter from ipad, 120  from  twitter web, 762 from twitter android and 628 from twitter iPhone.  

```{r}
table(tweets$statusSource)
```

\newpage
1c. We're going to compare the language used between the Android and iPhone sources, so only want to keep tweets coming from those sources.  Explain what the `extract` function (from the `tidyverse` package) is doing below.  In your explanation, be sure to include the purpose of each of the arguments (`col`, `into`, `regex`, `remove`).

> ANSWER: Extract is acting to create a new column (using into to name it) that says what the source  is. Regex is the filter for what to remove. Col says which column you are extracting from.  

```{r}
tweets2 <- tweets %>%
  extract(col = statusSource, into = "source"
          , regex = "Twitter for (.*)<"
          , remove = FALSE) %>%
  filter(source %in% c("Android", "iPhone"))
 
```


\newpage
## How does the language of the tweets differ by source?  

2a. Create a word cloud for the top 50 words used in tweets sent from the Android.  Create a second word cloud for the top 50 words used in tweets sent from the iPhone.  How do these word clouds compare?  (Are there some common words frequently used from both sources? Are the most common words different between the sources?)

Don't forget to remove stop words before creating the word cloud.

> ANSWER: They seem pretty similar, but some of the liberal words seem more prominent on iPhones, such as hillary and cnn while android users mentioned more conservative terms.   

```{r}
tweetsI <- tweets2%>%filter(source == "iPhone")
tweetsA <- tweets2%>%filter(source == "Android")

IPhone_word_frequencies <- tweetsI %>%
  unnest_tokens(output = word, input = text) %>%
  anti_join(stop_words, by="word") %>%
  count(word, sort = TRUE) 

Android_word_frequencies <- tweetsA %>%
  unnest_tokens(output = word, input = text) %>%
  anti_join(stop_words, by="word") %>%
  count(word, sort = TRUE)

IP <- IPhone_word_frequencies %>%
  with(wordcloud(word, n, max.words=50))

And <- Android_word_frequencies %>%
  with(wordcloud(word, n, max.words=50))


IP 
# And
```




\newpage
2b. Create a visualization that compares the top 10 *bigrams* appearing in tweets by each source (that is, facet by source).  After creating a dataset with one row per bigram, you should remove any rows that contain a stop word within the bigram.  

How do the top used bigrams compare between the two sources?

> ANSWER: Pretty similaraly. iPhone users seem to have more liberal terrms together (Hillary Clinton or Bernie Sanders) than the Republicans. 

```{r}
iPhoneBigram <- tweetsI %>% 
  unnest_tokens(output = word, input = text, token = "ngrams", n=2)%>%
  anti_join(stop_words, by="word")%>%
  count(word, sort = TRUE)



andBigram <- tweetsA %>%
  unnest_tokens(output = word, input = text, token = "ngrams", n=2)%>%
  anti_join(stop_words, by="word")%>%
  count(word, sort = TRUE)%>% filter()

IP <- iPhoneBigram %>%
  with(wordcloud(word, n, max.words=50))

And <- andBigram %>%
  with(wordcloud(word, n, max.words=50))

IP
And
```


\newpage
2c. Consider the sentiment.  Compute the proportion of words among the tweets within each source classified as "angry" and the proportion of words classified as "joy"  based on the NRC lexicon.  How does the proportion of "angry" and "joy" words compare between the two sources?  What about "positive" and "negative" words?  

> ANSWER: There is a pretty equal split on the Iphone  (176 joy 170 angry) but far more angry on the Android.  Therre are 794 positive  and  662 negaative on the Android and 449 positive and 266 negative on the Iphone. 

```{r}
ASent <- tweetsA%>%
  unnest_tokens(output = word, input = text)

nrc_And <- ASent %>%
  inner_join(nrc_lexicon, by = "word") %>%
  filter(sentiment == "anger" |  sentiment == "joy")%>%
  count(sentiment)

ISent <- tweetsI%>%
  unnest_tokens(output = word, input = text)

nrc_IP <- ISent %>%
  inner_join(nrc_lexicon, by = "word") %>%
  filter(sentiment == "anger" |  sentiment == "joy")%>%
  count(sentiment)

nrc_IP2 <- ISent %>%
  inner_join(nrc_lexicon, by = "word") %>%
  count(sentiment)

nrc_And2 <- ASent %>%
  inner_join(nrc_lexicon, by = "word") %>%
  count(sentiment)



```


\newpage
2d. Lastly, based on your responses above, do you think there is evidence to support Robinson's claim that Trump only writes the (angrier) Android half of the tweets from realDonaldTrump?  In 2-4 sentences, please explain.

> ANSWER: Yes I agree, as the iPhone seemed to have a substantally more positive and less angry ratio  than the Android.
